---
title: "Document_creation"
output: html_document
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE,    # Hides the code
  message = FALSE, # Hides package loading messages
  warning = FALSE# Hides warnings 
)
```
---
title: "Load Library"
output: html_document
date: "2026-04-25"
---
```{r load-data, message = FALSE}
knitr::opts_chunk$set(cache = T)## Keeps the document fast
library(pagedown)
library(readxl)
library(gt)
library(dplyr)
library(tidyr)
library(janitor)
library(gt)
library(ggplot2)
library(plotly)
library(lubridate)
library(tinytex)
```

  
# Crime Incident Report Data

```{r}
setwd("~/Documents/uni_UDE/AdvanceR/")
crime_data <- readRDS("Crime_Large.rds")
field_explanation <- readxl::read_xlsx("FieldExplanation.xlsx")


crime_data %>%
  head(10) %>%
  gt() %>%
  tab_header(
    title = "Crime Incident Reports Sample",
    subtitle = "Displaying the first 10 records from the dataset"
  ) %>%
  tab_options(
    table.width = pct(100)
  )
```

The "Crime Incident Reports" dataset contains `r nrow(crime_data)` individual observations 
across `r ncol(crime_data)` different variables. Based on the field explanations, 
the data provides detailed records of incidents ranging from `r min(crime_data$YEAR, na.rm = TRUE)` 
to `r max(crime_data$YEAR, na.rm = TRUE)`.

<span style="font-size: 0.8em;">The "Crime Incident Reports" dataset contains `r nrow(crime_data)` individual observations 
across `r ncol(crime_data)` different variables. Based on the field explanations, 
the data provides detailed records of incidents ranging from `r min(crime_data$YEAR, na.rm = TRUE)` 
to `r max(crime_data$YEAR, na.rm = TRUE)`.</span>


# Data Cleaning
```{r, message=FALSE, warning=FALSE}

# 1. Tidy the data
field_explanation_tidy <- field_explanation %>%
  clean_names() %>%
  fill(everything(), .direction = "down") %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  distinct() %>%
  # Splitting the column into three
  # Check if your data uses "_" or " ". If it's spaces, use sep = " "
  separate(
    col = field_name_data_type_required, 
    into = c("field_name", "data_type", "required"), 
    sep = "," 
  )
# 2. Display as a formatted table
field_explanation_tidy %>%
  gt() %>%
  tab_header(
    title = "Field Explanations: Tidy View",
    subtitle = "Definitions and metadata for the Crime Incident dataset"
  ) %>%
  tab_options(table.width = pct(100))
```

# Explorative Data Analysis

```{r eda-days, message=FALSE, warning=FALSE}
# 1. Prepare the data
# We group by day_of_week and count the number of rows (incidents)
day_counts <- crime_data %>%
  group_by(DAY_OF_WEEK) %>%
  summarise(incident_count = n()) %>%
  # Ensure days are in a logical order instead of alphabetical
  mutate(day_of_week = factor(DAY_OF_WEEK, 
                               levels = c("Monday", "Tuesday", "Wednesday", 
                                          "Thursday", "Friday", "Saturday", "Sunday")))

# 2. Create the Bar Chart
ggplot(day_counts, aes(x = DAY_OF_WEEK, y = incident_count, fill = day_of_week)) +
  geom_col() +
  scale_fill_brewer(palette = "Set3") + # Nice soft colors
  theme_minimal() +
  labs(
    title = "Crime Incidents by Day of the Week",
    x = "Day",
    y = "Total Incidents"
  ) +
  theme(legend.position = "none")

# 3. Create the Summary Table
day_counts %>%
  gt() %>%
  tab_header(
    title = "Frequency of Incidents",
    subtitle = "Summary table by day of the week"
  ) %>%
  cols_label(
    day_of_week = "Day",
    incident_count = "Number of Incidents"
  ) %>%
  fmt_number(columns = incident_count, decimals = 0) %>%
  tab_caption(caption = "Table 1: Total crime incidents recorded for each day of the week.")
# 1. Compute incidents by Year and Month and create the Date variable
time_counts <- crime_data %>%
  group_by(YEAR, MONTH) %>%
  summarise(incident_count = n(), .groups = "drop") %>%
  # Create a date object (e.g., 2024-01-01)
  mutate(Date = make_date(year = YEAR, month = MONTH, day = 1)) %>%
  arrange(Date)

# 2. Create the ggplot
# We use a line and points to show the trend and individual monthly totals
p <- ggplot(time_counts, aes(x = Date, y = incident_count, group = 1)) +
  geom_line(color = "#0072B2", linewidth = 1) +
  geom_point(color = "#0072B2", size = 1.5) +
  theme_minimal() +
  labs(
    title = "Crime Incident Trends Over Time",
    x = "Timeline",
    y = "Number of Incidents"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months")

# 3. Make the plot interactive
# ggplotly converts the static plot 'p' into an interactive version
ggplotly(p)
```
<ol>
  <li>Phase One
    <ol>
      <li>Sub-task A</li>
      <li>Sub-task B</li>
    </ol>
  </li>
  <li>Phase Two</li>
</ol>

<style>
ol { counter-reset: item; }
ol li { display: block; }
ol li:before { content: counters(item, ".") ". "; counter-increment: item; }
</style>

```{r plotly-filter, message=FALSE, warning=FALSE}
library(plotly)
library(crosstalk)
library(dplyr)
library(lubridate)

# 1. Prepare data with DISTRICT included
time_district_counts <- crime_data %>%
  group_by(YEAR, MONTH, DISTRICT) %>%
  summarise(incident_count = n(), .groups = "drop") %>%
  mutate(Date = make_date(year = YEAR, month = MONTH, day = 1)) %>%
  filter(!is.na(DISTRICT)) # Clean up any missing district info

# 2. Create the SharedData object
shared_crime <- SharedData$new(time_district_counts)

# 3. Create the filter widget
# This creates the dropdown menu
filter_dropdown <- filter_select(
  id = "district-selector",
  label = "Select a District",
  sharedData = shared_crime,
  group = ~DISTRICT
)

# 4. Create the plot using the SharedData
p_filtered <- ggplot(shared_crime, aes(x = Date, y = incident_count, color = DISTRICT, group = DISTRICT)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  theme_minimal() +
  labs(
    title = "Monthly Incidents by District",
    x = "Timeline",
    y = "Incident Count"
  ) +
  theme(legend.position = "none") # Dropdown replaces need for a messy legend

# 5. Arrange them together
# We use bscols to put the filter on top of the plot
bscols(
  widths = c(12),
  filter_dropdown,
  ggplotly(p_filtered, tooltip = c("Date", "incident_count", "DISTRICT"))
)
```

